\subsection{Problem and data}
The dataset originates in the cost-accuracy study of \citet{dehoop2022cost}
and is the one distributed with \citet{batlle2024kernel} and
\citet{mora2025operator}. An isotropic elastic plate occupies
$D=(0,1)^2$; the displacement field $w$ satisfies
$\nabla\cdot\sigma=0$ with a fixed constitutive law, displacement conditions
on the bottom and lateral parts of the boundary, and a prescribed normal
traction $\bar t$ on the top edge $\Gamma_t=[0,1]\times\{1\}$. The quantity of
interest is the von Mises stress field $\sigma_v$ on $D$. The learning task is
the map $G:\bar t\mapsto\sigma_v$. Loads are drawn from the Gaussian field
$\mathcal{GP}\big(100,\,400^2(-\Delta+3^2 I)^{-1}\big)$ with homogeneous
Neumann boundary conditions for the Laplacian; outputs are finite element
solutions interpolated to a regular $41\times41$ grid, and the load is sampled
at $41$ points. The distributed file contains 40000 input/output pairs; the
input array stores the load broadcast along the second grid coordinate, which
we verified is an exact copy (maximal deviation $0$ across all 40000
samples), so all our methods consume the load as a vector in $\mathbb R^{41}$.

Errors are mean relative $L^2$ errors over the test set,
\[
\frac1{N_{\mathrm{test}}}\sum_{n}\frac{\|\widehat
G(u_n)-G(u_n)\|_2}{\|G(u_n)\|_2},
\]
computed on grid values in double precision. Quadrature weighting
(trapezoidal instead of plain vector norms) changes our reported numbers by
less than 0.02 percentage points, and we report the plain-norm convention of
the prior work.

\subsection{Protocols and published results}
Two regimes appear in the literature, and we follow both. In the
\emph{high-data} protocol the first 20000 samples are available for training
and the last 20000 form the test set; \citet{dehoop2022cost} report, at 20000
training samples, 4.55\% (PARA-Net), 4.67\% (PCA-Net), 4.76\% (FNO) and 5.20\%
(DeepONet), and \citet{batlle2024kernel} report 5.18\% for their
optimal-recovery kernel method (Mat\'ern/rational quadratic; 27.11\% for the
linear kernel) on the same task. Within the 20000-sample budget we hold out
the last 1000 samples (of a fixed permutation) for model selection and
stacking, and train on the remaining 19000, so no method of ours sees more
than the 20000 training samples used by the baselines. In the \emph{low-data}
protocol of \citet{mora2025operator}, 1250 samples are available and the same
20000-sample test set is used; their table gives 8.70\% (DeepONet), 6.62\%
(FNO), 6.95\% (the kernel method of \citealt{batlle2024kernel}), 6.74\% (their
zero-mean GP), 7.12\% (DeepONet-mean GP) and 6.49\% (FNO-mean GP). In this
regime we carve the validation split (250 samples) out of the 1250, so
training uses 1000 samples and every choice made by the pipeline is informed
by the 1250 available labels only.

\subsection{A data-driven check of the mirror symmetry}
\label{sec:symcheck}
The continuous problem is invariant under $x_1\mapsto1-x_1$: the domain and
boundary partition are symmetric, and the input law has a symmetric
covariance. On the grid, reflecting the load ($S$) should reflect the stress
field along the first grid coordinate ($T$), i.e.\ $G(Su)=TG(u)$. Rather than
assume this, we test it on the data. For each of the first 200 samples we
searched all 40000 loads for the nearest neighbor of the reflected load $Su_i$
and compared the corresponding outputs. For the five best-matching pairs, the
input mismatch $\|Su_i-u_j\|/\|u_j\|$ ranges over $0.27$--$0.32$, while the
output mismatch $\|Tv_i-v_j\|/\|v_j\|$ ranges over $0.085$--$0.14$; reflecting
along the second coordinate instead gives output mismatches of order one.
Outputs of near-mirror inputs are far closer than the inputs themselves,
which is what equivariance plus a Lipschitz solution map predicts, and the
effect singles out the correct output reflection axis. A complementary check
on the input law: the empirical mean and covariance of the training loads are
invariant under reflection to within 1.1\% and 1.5\% respectively, as required
for Proposition~\ref{prop:sym}. Consistently with all of this, reflection
averaging at test time improves every model we trained
(Section~\ref{sec:experiments}), as Proposition~\ref{prop:sym} says it must on
average when the symmetry holds.

\subsection{Related work}
\label{sec:related}
The benchmark sits at the meeting point of three lines of work. \emph{Neural
operators} learn maps between function spaces: DeepONet \citep{lu2021deeponet}
with its branch/trunk factorization, the Fourier neural operator
\citep{li2021fourier} and the broader neural-operator framework
\citep{kovachki2023neural}, and the reduced-basis PCA-Net and PARA-Net
architectures assembled for the cost--accuracy study of
\citet{dehoop2022cost}. These methods are fast and mesh-flexible but do not by
themselves come with error control, and it is their numbers on this problem
that define the plateau we start from. \emph{Kernel and Gaussian process
methods} for operator learning approach the same maps through reproducing
kernels: the optimal-recovery framework of \citet{batlle2024kernel}, with its
convergence theory and a-priori bounds, is the most direct comparison, and it
is competitive with neural operators on most of the benchmarks it considers.
Data-adapted kernels learned by cross-validation \citep{darcy2023one} and
vector-valued kernel formulations extend the reach of this family. \emph{Hybrids}
that place a neural network and a kernel in the same estimator are the closest
relatives of our pipeline: \citet{mora2025operator} use a neural operator as
the mean of a Gaussian process and fit the two together, and this is the work
we most directly build on, differing in that we fit the mean first and the
kernel after, tune by cross-validation rather than marginal likelihood, and
solve the correction exactly at nineteen thousand points.

Two further threads inform the design. The kernel-flow method
\citep{owhadi2019kernel} learns a kernel from data by minimizing a
half-sample cross-validation loss, and its use as a regularizer for the inner
layers of a network \citep{yoo2021deep} is exactly the term we analyze in
Proposition~\ref{prop:kf}; kernel flows have since been used at scale as
emulators for physical forward models, for instance in atmospheric radiative
transfer retrievals \citep{susiluoto2025radiative} and in the inference of
convective-storm structure from satellite observations
\citep{prasanth2021kernel}, where presenting the input in its physical
parametrization and adapting the kernel to data are what make the emulator
accurate. Our pipeline follows the same instinct, with the smoothing of the
target performed by a neural ensemble rather than by a hand-chosen reduction.
Finally, the effective-dimension reading of the kernel solve
(Lemma~\ref{lem:deff}) draws on the random-matrix description of kernel Gram
spectra \citep{koltchinskii2000random} and on the classical
Mar\v{c}enko--Pastur \citep{marchenko1967distribution} and spiked-covariance
\citep{baik2005phase} laws; the same effective dimension governs the
generalization of kernel ridge regression \citep{caponnetto2007optimal}.
