\paragraph{What moved the number, and what stopped it.}
The improvement over the published band decomposes unevenly. Accurate neural
means trained on the reported metric and averaged over the reflection
symmetry do most of the work; stacking members from different architecture
families adds exactly what their measured residual correlations permit, about
a tenth of a point here; the kernel corrections contribute a few hundredths
more and, with them, the certificate of Theorem~\ref{thm:or} and the
calibrated band of Proposition~\ref{prop:conf}. We did not find evidence that
any published architecture was mis-tuned by its authors; the components
compose, and the composition had not been tried on this benchmark. What stops
further movement is not a missing architecture. Section~\ref{sec:diversity}
locates the remaining error in a component that all families share, that
concentrates where the finite element data are least reliable, that no input
statistic predicts, and that neither capacity, nor diversity, nor more data
reduces. The margin over the published methods other than PARA-Net is real,
and the match with PARA-Net is exact to within run-to-run noise, but both
should be read for what they are: the last few learnable hundredths above a
floor set by the data, not a step on the way to zero.

\paragraph{Relation to neural-mean Gaussian processes.}
Our correction stage is closest in spirit to \citet{mora2025operator}, who
place a neural operator inside the mean of a Gaussian process and fit both
jointly. The differences are operational but they matter at this scale: we
fit the mean first and the kernel after (so the expensive stage is ordinary
network training), we tune the kernel by validation rather than by marginal
likelihood, we correct with an exact solve at $n=19000$ rather than train
through the kernel, and we add the feature-space second stage. The theory in
Section~\ref{sec:theory} applies verbatim to their setting as well; the
measured drop in the fitted RKHS norm (Table~\ref{tab:norms}) is the
quantitative reason residual corrections of accurate means are the right
place to spend kernel capacity.

\paragraph{Relation to kernel emulation practice.}
The pipeline is consistent with the experience of the kernel-flow line of
work in emulation at scale. The closest instance is the forward-model
emulator of \citet{susiluoto2025radiative} for the OCO-2 CO$_2$ retrievals,
in which a Gaussian process with a cross-validation-learned kernel replaces
the full-physics radiative transfer code within measurement-error precision;
kernel flows have likewise been used to infer convective-storm structure
from passive microwave observations \citep{prasanth2021kernel}. The lesson of
that work matches ours: kernels with data-adapted hyperparameters are
extremely effective once the input is presented in its physical
parametrization and the target has been reduced to something smooth. Here the
reduction is performed by the neural ensemble rather than by physics, and the
kernel-flow loss itself admits the leave-half-out reading of
Proposition~\ref{prop:kf}. The OCO-2 setting is also the natural next test
bed for the residual coupling studied here. Its training pairs are
simulator-generated state-to-radiance maps of exactly the shape this paper
exploits, a low-dimensional physical state mapped to a smooth
high-dimensional output, the mission's Level 1 and Level 2 products are
publicly distributed through NASA's Earthdata archive, and the emulation
setup of \citet{susiluoto2025radiative} specifies the state parametrization
and sampling design, so a neural mean with a validated kernel correction and
a conformal wrapper can be evaluated there against the pure-kernel emulator
without new data collection.

\paragraph{Limitations.}
The benchmark has a one-dimensional input function and a fixed geometry; the
exact kernel solve exploits both, and problems with high-dimensional input
fields would require the usual approximations (inducing points, random
features) at some cost to the certificates. The certified bound of
Theorem~\ref{thm:or} controls the error relative to the RKHS norm of the
residual operator, a quantity we can only lower-bound empirically; the
conformal calibration we report is the honest, assumption-light complement.
Training the means on the metric, the reflection steps, and the stacking are
benchmark-agnostic, but the specific error levels reported here should be
read as properties of this dataset and protocol. Finally, our low-data
protocol reproduces that of \citet{mora2025operator} up to the unavoidable
ambiguity of which 1250 samples are used; we use the first 1250 of the
training block and report the same 20000-sample test set, and we release
splits and code so the comparison can be audited.

\paragraph{What the second problem settles.}
The OCO-2 study keeps every component and inverts the balance, and the pair
of problems brackets the design space. When the network and the kernel tie
(structural mechanics), the coupling pays: stack them, correct the residual,
and the certified and conformal machinery rides along. When the kernel
trails by an order of magnitude (OCO-2), the coupling is not the point; the
kernel's value moves inside the network, as an exact head on its learned
features, and the pipeline's job becomes metric management, training each
member in the metric it will be scored in and combining per coordinate. In
both regimes the same two measurements decide everything in advance: the
pairwise residual correlations, which set the ensembling floor of
Corollary~\ref{cor:floor} at whatever level the members share their errors,
and the member error ratio, which the second-moment identity turns into a
keep-or-drop verdict for each candidate. Nothing in that protocol is
specific to these two datasets.

\paragraph{Outlook.}
Three directions seem worth pursuing. First, the benchmark itself:
regenerating the dataset with refined meshes at the corner singularities,
higher-order elements, or output grids that resolve the concentrations would
move the floor that Section~\ref{sec:diversity} measures, and would return
the problem to being a test of surrogates rather than of its own data. The
diagnostic protocol used there, cross-family residual correlation, the
second-moment prediction of Proposition~\ref{prop:secmom}, and the
scaling-in-$n$ check, is cheap to run on any benchmark suspected of the same
condition. Second, the same recipe on the remaining benchmarks of
\citet{dehoop2022cost} and \citet{batlle2024kernel}, where the input
functions are genuinely high-dimensional and the interplay between neural
means and kernel corrections should look different. Third, the uncertainty
side: $P_\lambda$ is a design quantity, so it can be optimized, and the
connection of Lemma~\ref{lem:deff} between the nugget, the spectrum, and the
effective dimension suggests principled ways to spend a fixed computational
budget on the correction stage.
