All experiments ran on a single laptop (one NVIDIA RTX~PRO~2000, 8~GB; 16 CPU
cores, 64~GB RAM), in single precision for network training and double
precision for all kernel solves and error computations. The dataset is the
distributed \texttt{StructuralMechanics} file pair (40000 samples); inputs are
reduced to $\mathbb R^{41}$ after verifying the broadcast structure exactly.
Splits: the training block is samples 1--20000, the test set is samples
20001--40000. High-data protocol: a fixed permutation of the training block
reserves 1000 samples for validation, 19000 for fitting. Low-data protocol:
the first 1250 samples of the training block, of which the last 250 form the
validation split. The test set is never touched during development; each
configuration is evaluated on it once, after selection on validation.

\paragraph{FNO.} Width 64, 14 modes per dimension, 4 spectral layers with
pointwise linear skips and GELU, lift from 3 channels (broadcast load, two
coordinates), projection $64\to128\to1$. AdamW, learning rate
$1.5\times10^{-3}$ ($2\times10^{-3}$ with batch 256), weight decay $10^{-6}$,
cosine schedule to $10^{-6}$, 300 epochs, batch 256. 6.45M parameters.

\paragraph{Transformer (implemented, not in the reported ensemble).} Encoder:
41 tokens (load value and 10 sine/cosine pairs of the coordinate), 5 pre-norm
blocks, dimension 192, 4 heads, MLP ratio 4. Decoder: grid queries from
Fourier features of both coordinates, two cross-attention blocks, pointwise
head $192\to192\to1$; 3.16M parameters. The cross-attention over 1681 grid
queries is the most compute-intensive of our means, and the hardware available
for this study (a single laptop GPU that proved unstable under sustained load)
did not let us train it to the level of the other members; we therefore
exclude it from the reported ensemble and note it as the natural fourth
architecture family for a follow-up on stable hardware.

\paragraph{UNet.} Three convolutional scales ($41\to21\to11$ by average
pooling, ceil mode) and a bottleneck at $6\times6$; two $3\times3$
convolutions with GroupNorm(8) and SiLU per block; widths $48/96/192/384$;
bilinear upsampling with skip concatenation; $1\times1$ output head. Input
channels as for the FNO. AdamW, learning rate $1.5\times10^{-3}$, weight
decay $10^{-5}$, cosine schedule, 200 epochs, batch 256. 4.38M parameters.

\paragraph{MSE variant.} The residual MLP trained with mean squared error on
standardized targets in place of the metric loss, otherwise identical recipe;
it enters the ensemble as a loss-diversity member and is also the strongest
plain MLP we obtain.

\paragraph{MLP.} Input layer $41\to1024$, three residual SiLU blocks of width
1024, output $1024\to1681$. AdamW, learning rate $10^{-3}$, weight decay
$10^{-5}$, cosine schedule, 400 epochs, batch 256. 4.91M parameters. A wider
variant (width 1536, five residual blocks, 14.5M parameters) is trained under
the same recipe.

\paragraph{Refiner.} The same residual trunk with input layer
$(41+1681)\to1024$: the load is concatenated with the kernel method's stress
prediction for that load. Training uses four-fold out-of-fold kernel
predictions for the input channel and full-data kernel predictions at
evaluation; otherwise identical to the MLP. 6.64M parameters. Reflection
augmentation flips the load and permutes both the kernel channel and the target
by the row-reversal index.

\paragraph{Common training details.} Loss: mean over the batch of
$\|\hat v-v\|_2/\|v\|_2$ in original units (predictions denormalized inside
the loss; targets standardized by the per-pixel training mean and global
standard deviation), except for the MSE variant above. Reflection
augmentation with probability $1/2$; reflection averaging at evaluation.
Model selection: validation error computed every 10 epochs, best checkpoint
kept. A single seed is trained per architecture; the ensemble's diversity
comes from the architectures and losses rather than from reseeding, and the
correlation analysis of Section~\ref{sec:diversity} indicates same-architecture
reseeds would be more correlated still.

\paragraph{Stacking.} Global weights: convex, initialized at the simplex
minimizer of the measured second-moment matrix $S$
(Proposition~\ref{prop:secmom}) and polished by a short random search on the
validation metric. Per-pixel weights: affine in the members at each grid
point, ridge parameter $10^{-3}$, fitted on half the validation split and
accepted only if they beat the global weights on the held-out half, then
refitted on the full split.

\paragraph{Kernel stages.} Mat\'ern-$5/2$ on standardized inputs. Scale grid
$\{0.5,1,2,4\}\times$ the median pairwise distance (estimated on 2000
points), nugget grid $\{10^{-7},10^{-5},10^{-3}\}$ (scaled by $n$), tuned on
validation using an 8000-sample subsample of the training set, refit at the
chosen pair on the full training set in double precision. The pure kernel
baseline (Table~\ref{tab:main}) uses the grid
$\{0.5,0.75,1,1.5,2\}\times$median and nuggets $\{10^{-8},10^{-6},10^{-4}\}$
directly at $n=19000$. Feature stage: penultimate activations of the ensemble
members (FNO: spatially averaged pre-projection channels, 128; transformer:
mean-pooled encoder tokens, 192; MLP: trunk output, 1024), concatenated,
standardized, same kernel family and tuning.

\paragraph{Uncertainty.} Posterior standard deviation \eqref{eq:power}
computed from the Cholesky factor of $K+n\lambda I$ by triangular solves.
Conformal scaling: the 0.9-quantile of $\|e\|_2/P_\lambda$ on the validation
split multiplies $P_\lambda$ on test; coverage is the fraction of test
samples whose absolute error norm falls below the scaled band.
