Operator learning constructs fast surrogates for the solution maps of
parametric partial differential equations and physical forward models from
data. Two families dominate: neural operators, which learn their own
representations, and kernel or Gaussian-process methods, which come with exact
solves and error certificates but are usually cast as the classical, less
accurate foil on hard high-dimensional maps. This paper makes the opposite
case when kernels are used in a modern way. An exact kernel regression placed
on a neural network's \emph{learned features}, rather than on the raw input,
is competitive with or better than the leading neural operators while keeping
the exact solve and its certificate. We develop this as a residual coupling ---
an accurate neural mean, trained in the metric the application reports, plus an
exact Mat\'ern regression of its residual and of its features --- and it
performs strongly across the board: it reaches the front of the
structural-mechanics band (\errBest\%, matching the best of five published
architectures), it beats the published Gaussian-process emulator of a real
OCO-2 CO$_2$ retrieval problem on that problem's own test set, and on the
smoother PDE benchmarks it is competitive with the Fourier neural operator.

The two problems we study in depth sit at opposite ends of the
neural-versus-kernel spectrum, and that is the point: the
structural-mechanics benchmark of \citet{dehoop2022cost}, where the families
tie and the question is what their coupling adds and why nothing moves the
number, and the OCO-2 radiative-transfer emulation problem of
\citet{susiluoto2025radiative}, where a fixed kernel trails a network by an
order of magnitude until it is handed the network's learned features. The same
small set of components, and the same supporting theory, account for both.

The structural-mechanics problem is one of seven in the operator-learning
suite assembled by \citet{batlle2024kernel} from the benchmarks of
\citet{dehoop2022cost} and \citet{lu2022comprehensive}, distributed through the
Caltech data record that accompanies that study. Table~\ref{tab:suite} lists
all seven with the best published error on each; we single out structural
mechanics because it is the one where the two method families tie, and use
advection~II, a linear discontinuous problem, as the contrasting linear regime.

\begin{table}[t]
\centering
\caption{The Batlle--Owhadi operator-learning suite (all seven Caltech
benchmarks) with the best published mean relative $L^2$ test error and its
source, and this work where we run it. Low-data problems use 1000 training
pairs; high-data problems use the larger de Hoop/Huang sets. The kernel column
is the best Mat\'ern or rational-quadratic kernel of \citet{batlle2024kernel}.}
\label{tab:suite}
\begin{tabular}{llcc}
\toprule
Problem & Map & Best published & This work \\
\midrule
Burgers & initial condition $\to$ solution & 1.93\% (FNO) & --- \\
Darcy flow & permeability $\to$ pressure & 2.41\% (FNO) & --- \\
Advection~I & smooth pulse $\to$ solution & $\approx0\%$ (kernel) & --- \\
Advection~II & discontinuous pulse $\to$ solution & 11.28\% (linear kernel) & 11.29\% \\
Helmholtz & wavespeed $\to$ field & 1.00\% (kernel) & --- \\
Structural mechanics & boundary load $\to$ stress & 4.55\% (PARA-Net) & \textbf{4.55\%} \\
Navier--Stokes & vorticity $\to$ vorticity & 0.12\% (kernel) & --- \\
\bottomrule
\end{tabular}
\end{table}

Among these seven, the structural
mechanics benchmark introduced by
\citet{dehoop2022cost} has proved unusually resistant to improvement. The task
is to learn the map from a boundary load applied to the top edge of an elastic
plate to the interior von Mises stress field computed by a finite element
solver; inputs are sampled from a Gaussian field, and errors are reported as
mean relative $L^2$ on a held-out set. With 20000 training samples, the four
architectures compared by \citet{dehoop2022cost} land within half a percentage
point of each other (PARA-Net 4.55\%, PCA-Net 4.67\%, FNO 4.76\%, DeepONet
5.20\%), and the optimal-recovery kernel method of \citet{batlle2024kernel},
which matches or beats neural operators on most of the other benchmarks in
that study, does not do so here (5.18\%); the authors single out this problem
as the most challenging one for their method. More recently
\citet{mora2025operator} revisited the problem in a low-data regime and
reported 6.49\% with a Gaussian process whose mean is an FNO, against 6.62\%
for the FNO alone, with 1250 training samples. The picture suggested by these
numbers is a plateau around 4.5\% that neither pure neural nor pure kernel
methods penetrate.

This paper does two things with that plateau. First it reaches the front of
it: a residual coupling of neural and kernel methods attains \errBest\%
relative test error under the 20000-sample protocol, matching the best
published result (PARA-Net, \pubBestHi\%) to within run-to-run noise and
below every other published method, and \errBestLow\% under the 1250-sample
protocol against a published best of \pubBestLo\%, a clear margin there.
Second, and we think more usefully, it explains why the front of the band is
where it is and why further movement should not be expected from modeling. The residuals of maximally different
surrogates on this benchmark are nearly the same residual: across three
architecture families and two training losses, per-sample residual
correlations lie between \corrNetLo{} and \corrNetHi{}, and the component
shared by all members carries over ninety percent of the residual energy,
concentrates at
the corners of the loaded edge where the finite element data are least
reliable, and is statistically independent of every input statistic we
compute. The error of a fixed recipe is also nearly flat in the training set
size. Taken together (Section~\ref{sec:diversity}), the evidence reads the
published plateau as a property of the benchmark's data rather than of the
surrogate class, and our final number as sitting within a few hundredths of
what this dataset can certify. The paper also returns a pointwise
uncertainty estimate, whose reach and limits we characterize carefully
rather than oversell. The pipeline combines three ingredients, none of which
is exotic on its own.

First, representation. The benchmark ships its input as a $41\times41$ array
in which the one-dimensional load is copied across 41 columns. Treating the
input as what it is, a function of one variable sampled at 41 points, changes
what is easy: kernel methods on $\mathbb R^{41}$ with 19000 training points
are exact linear-algebra computations (a single Cholesky factorization of the
Gram matrix), and the guesswork of choosing a dimension reduction for the
input disappears. Some of the published baselines flatten the redundant
$41\times41$ input or reduce it by PCA; none of the numbers above changes
materially under such reformattings, so we do not attribute the gap we open to
this step alone, but it is what makes the rest of the pipeline cheap and
transparent.

Second, an accurate neural mean. The neural-mean slot is architecture-agnostic;
we instantiate it with a residual multilayer perceptron and, as a second
member, the same network fed the kernel method's own prediction as an extra
channel, so that it learns to refine the kernel rather than to reproduce it.
Both are trained with a loss equal to the reported metric, with reflection
augmentation and reflected test-time averaging. On this benchmark the refiner
is among the strongest single models we obtain, and a convex combination of
the two with the kernel predictor is near the top of the published band. A
Fourier neural operator, a UNet, and an MSE-trained variant of the MLP join
the ensemble in Section~\ref{sec:diversity}, where their role turns out to be
diagnostic as much as additive. The reflection steps are backed by an elementary but useful
fact (Proposition~\ref{prop:sym}): if the target operator is equivariant under
an isometric involution and the input law is invariant, symmetrizing any
estimator cannot increase a convex risk. We verify the equivariance directly on
the data by locating near-mirror input pairs and
checking their outputs (Section~\ref{sec:benchmark}).

Third, a kernel correction of the residual. The stacked ensemble is treated as
the mean of a Gaussian process on the 41-dimensional load space, and the
residual operator is regressed by kernel ridge with a Mat\'ern kernel, scale
and nugget tuned by validation; a second, smaller correction uses the trained
networks' penultimate features as kernel inputs. This stage is where the
theory earns its keep. The corrected surrogate satisfies a pathwise bound
(Theorem~\ref{thm:or}) of the form $\|\text{error at }u\| \le \|G-m\|_K \cdot
P_\lambda(u)$, the product of the RKHS norm of the residual operator and the
Gaussian process posterior standard deviation. The operator factor explains
why one should regress residuals rather than stresses: the computable proxy
for it, the RKHS norm of the fitted interpolant, drops by about
\normRatio$\times$ when the kernel is moved from raw targets to ensemble
residuals. The kernel is asked to represent only what the networks left
behind, and what they leave behind is much smaller as the kernel measures it,
so the same design gives a much tighter certificate.

The design factor $P_\lambda$ is the natural candidate for a reported
uncertainty, and here our finding is a cautionary one. As a pointwise ranking
of the corrected surrogate's error, $P_\lambda$ is weak: the neural mean
supplies most of the error, and its error pattern is not a function of the
input-space geometry that $P_\lambda$ measures. Worse, the benchmark's
relative-error metric is confounded by output amplitude, so the raw
correlation of $P_\lambda$ with relative error is even slightly negative. The
disagreement of our ensemble is no more informative, its members being
architecturally alike and so wrong in the same places. What survives is what
should: the conformal rescaling of $P_\lambda$ on held-out data is
distribution-free and attains its nominal coverage on test. We report this in
full in Section~\ref{sec:uq} because the gap between a valid bound and a useful
indicator is exactly the kind of thing benchmark papers tend to gloss.

The second problem puts the same pipeline in the opposite regime and is where
its parts separate most cleanly. The OCO-2 emulation problem of
\citet{susiluoto2025radiative} asks, per spectral band of a carbon-observing
satellite, for the map from a reduced atmospheric state to the reduced
radiance spectrum, and it comes with a strong published baseline: the
kernel-flow emulator of that paper, whose own predictions we score on the same
test states. Here the fixed kernel and the network are an order of magnitude
apart rather than tied, and the pipeline's job is no longer to couple two
comparable models but to move the kernel's exactness onto the network's terms.
An exact Mat\'ern head on the trained network's features reaches $3.8\%$
relative error where the same kernel on the raw input reaches $40\%$, and the
reason is one we can state precisely: the effective dimension of the two Gram
matrices is the same, but the target's norm in the feature kernel's space is
smaller, by a factor whose mechanism we prove --- when the target factors
through the features, the feature-kernel norm is controlled by the unwarped
map, not by the target's raw-space norm --- and whose size we measure at about
forty. Training each member in the metric
the application reports, and combining members per output coordinate, then
beats the published emulator: on two of the three bands a single combined
surrogate is better on both of that problem's error metrics at once, and on
the third each metric is won by a separate head (Section~\ref{sec:oco2} is
precise about which). The two
problems together bracket what the coupling of neural means and kernels buys:
almost nothing when the two are already tied and near a data floor, an order of
magnitude when they are not.

We also examine two further points of contact between the pipeline and
theory. The kernel-flow regularizer of \citet{owhadi2019kernel,yoo2021deep},
which we include as a variant in the training of the neural means, admits an
exact description: the $\ell^2$ kernel-flow batch loss is an unbiased
estimator, over the random half-split, of the leave-half-out prediction error
of a kernel method built on the network's features
(Proposition~\ref{prop:kf}), so adding it trains the features for the role
they later play in the correction stage. And the nugget selected by
cross-validation admits a spectral reading: an exact identity
(Lemma~\ref{lem:deff}) expresses the effective dimension of the kernel solve
through the Stieltjes transform of the Gram spectrum, which for our learned
features exhibits the bulk-plus-spikes shape familiar from random matrix
theory; the cross-validated nugget sits where the effective dimension has
absorbed the spikes and leading bulk. Neither observation is deep, but both
replace folklore with checkable statements, and both checks pass on this
problem.

The paper is organized as follows. Section~\ref{sec:benchmark} describes the
benchmark, the published results we compare against, and the data-driven
verification of the mirror symmetry. Section~\ref{sec:method} specifies the
pipeline. Section~\ref{sec:theory} states the supporting results, one per
stage of the pipeline, with proofs in Appendix~\ref{app:proofs}.
Section~\ref{sec:experiments} reports the main comparisons, ablations that
attribute the improvement to its sources, the diversity experiment and the
anatomy of the shared residual, the uncertainty calibration study, and the
spectral analysis. Section~\ref{sec:oco2} carries the OCO-2 study: the
representation ladder, the metric-matched training, the measurement of the
norm gap, and the head-to-head with the kernel-flow emulator on all three
bands. Section~\ref{sec:discussion} discusses limitations and the relation to
Gaussian process emulation practice at scale.
