The pipeline has three stages: neural means, stacking, and kernel
corrections. All components operate on the load vector $u\in\mathbb R^{41}$
(standardized by training statistics) and produce stress fields in
$\mathbb R^{41\times41}$; all networks are trained with the reported metric as
the loss, i.e.\ the per-sample relative $L^2$ error in original units, which
we found mildly but consistently better than mean squared error on
normalized targets.

\subsection{Neural means}
\emph{Residual MLP.} The primary mean is a residual multilayer perceptron:
three or five residual SiLU blocks of width 1024--1536 mapping
$\mathbb R^{41}\to\mathbb R^{1681}$ directly. This is essentially PARA-Net
\citep{dehoop2022cost} with a modern training recipe, and it also supplies the
penultimate features used by the feature-space kernel stage.

\emph{Kernel-conditioned refiner.} The second mean is the same residual network
given an extra input channel: the kernel method's predicted stress field for
the same load, concatenated with the load. It therefore learns to correct the
kernel predictor rather than to reproduce the map from scratch, and it is the
strongest single model in our study. During training the refiner reads
\emph{out-of-fold} kernel predictions (four-fold, so the kernel channel is
never fit on the target sample), and at test time the full-data kernel
prediction; this keeps the channel honest.

\emph{Other instances.} The mean slot is not tied to a particular
architecture. We also use a Fourier neural operator \citep{li2021fourier}
that consumes the broadcast load as a field, a UNet on the same field
representation, and an MSE-trained variant of the MLP; all are drop-in means,
all enter the ensemble of Section~\ref{sec:diversity}, and their pairwise
error correlations are one of the measurements of this paper. We additionally
implement an encoder--decoder transformer
\citep{vaswani2017attention,dosovitskiy2021image} that tokenizes the 41 load
samples and decodes the field by cross-attention at the grid points; it is a
natural fourth architecture family, but we were unable to train it to the
level of the others on the hardware available for this study and report the
ensemble without it (Section~\ref{sec:diversity} returns to this).

All means are trained with AdamW and a cosine schedule, batches of 128--256,
for up to a few hundred epochs (Appendix~\ref{app:impl} lists every
hyperparameter).
During training each batch is reflected with probability $1/2$ (load reversed,
target field reflected along the first grid coordinate); at test time each
model is evaluated as $\frac12(f(u)+T f(Su))$. Proposition~\ref{prop:sym}
guarantees the test-time step cannot hurt on average, and the ablation
confirms small consistent gains from both steps. Optionally, the training
loss carries a kernel-flow term $\beta\,e_2$ computed from the batch
(Section~\ref{sec:theory-kf}) with a Gaussian kernel on pooled penultimate
features whose log-bandwidth is trained jointly; this variant appears in the
ablation.

\subsection{Stacking}
With $M$ trained models $f_1,\dots,f_M$ we form the convex combination
$m(u)=\sum_m w_m f_m(u)$, $w\in\Delta^{M-1}$, minimizing the validation
relative error; the weights are initialized at the simplex minimizer of the
measured second-moment matrix of Proposition~\ref{prop:secmom} and polished
by a short search on the 1000 held-out samples (250 in the low-data
protocol). A per-pixel variant allows the weights (plus an intercept) to
vary over the grid, ridge-fitted on half the validation split and accepted
only when it beats the global weights on the other half; it is the variant
the final pipeline uses. Stacking on a held-out split rather than uniform
averaging costs nothing, guards against a weak member
\citep{wolpert1992stacked}, and is statistically almost free at this size
(Proposition~\ref{prop:stack}).

\subsection{Kernel corrections}
\label{sec:method-kernel}
The stacked mean is then corrected by kernel ridge regression of its
residuals. Let $X=(u_1,\dots,u_n)$ be the training loads (standardized
coordinatewise), $R\in\mathbb R^{n\times1681}$ the matrix of training
residuals $v_i-m(u_i)$, and $k$ a Mat\'ern-$5/2$ kernel
$k(u,u')=\kappa\big(\|u-u'\|/s\big)$. The correction is
\[
\widehat r(u)=k(u,X)\,\alpha,\qquad \alpha=(K+n\lambda I)^{-1}R ,
\]
with $(s,\lambda)$ chosen on the validation split from a small grid ($s$ as a
multiple of the median pairwise distance, $\lambda\in\{10^{-7},10^{-5},10^{-3}\}$;
the grid is searched on an 8000-sample subsample and the chosen pair is refit
on all of $X$). The corrected surrogate is $m+\widehat r$. A second corrective
stage repeats the construction with the kernel evaluated on deep features,
the concatenated penultimate activations of the ensemble members
(reflection-averaged, standardized), fitted to the residuals of $m+\widehat
r$; it contributes a smaller improvement and is included when it helps on
validation. At $n=19000$ the exact solve is a single Cholesky factorization
of a $19000\times19000$ matrix, about a minute in double precision on a
laptop CPU, and prediction is two matrix products; no inducing points,
preconditioners, or stochastic solvers are involved. The Gaussian process
reading of the same formulas supplies the posterior standard deviation
$P_\lambda(u)$ of \eqref{eq:power} at negligible extra cost (one triangular
solve per evaluation batch), which is the uncertainty estimate studied in
Section~\ref{sec:experiments} and certified by Theorem~\ref{thm:or}.

Two remarks on scope. Nothing in the construction uses properties of this
particular PDE beyond the verified reflection symmetry, so the recipe
(native input parametrization, symmetrized accurate means, validated kernel
correction of the residual, posterior standard deviation as certificate)
transfers to other operator learning problems with low-dimensional input
parametrizations. And the pipeline degrades gracefully: dropping the feature
stage, the stacking, or the corrections recovers progressively simpler
methods whose individual numbers appear in the ablation table.
