The structural-mechanics benchmark put the neural mean and the kernel in the
balanced regime. To see the other regime we apply the same components to the
radiative-transfer emulation problem of \citet{susiluoto2025radiative}: per
spectral band of the OCO-2 instrument, learn the map from a reduced
atmospheric state ($20$--$24$ dimensions after the dimension reduction of
that paper) to the reduced radiance ($40$ PCA coefficients of the
monochromatic spectrum), from $20000$ training pairs. The data, and the
kernel-flow emulator's own predictions on the $2000$ test states, are public
(OSF project \texttt{u2t8a}), so the comparison below is computed on
identical test points against the published emulator itself rather than
against our reimplementation of it.

Two error metrics matter, and they disagree in an instructive way. The
\emph{reduced} metric is the relative $L^2$ error on the $40$ standardized
coefficients. The \emph{radiance} metric maps predictions back to the
monochromatic spectrum through the stored PCA projection and norms before
comparing; because the projection is orthogonal, on the reduced side it is
exactly the diagonal weighting $s_z$, and its energy concentrates almost
entirely on the first few coefficients.

\begin{table}[t]
\centering
\caption{OCO-2 O2 band, $2000$ held-out states, identical test points. The
kernel-flow row is the emulator of \citet{susiluoto2025radiative}, scored
from its own stored predictions. ``Features'' means the penultimate
activations of the trained network; the kernel head is the same exact
Mat\'ern solve as everywhere else in this paper.}
\label{tab:oco2}
\begin{tabular}{lcc}
\toprule
model & reduced & radiance \\
\midrule
Mat\'ern kernel, raw input, one length scale & 40.47\% & 0.240\% \\
Mat\'ern kernel, raw input, sensitivity-scaled (ARD) & 30.01\% & --- \\
kernel-flow emulator \citep{susiluoto2025radiative} & 16.89\% & 0.0448\% \\
residual MLP, flat metric & 3.99\% & 0.159\% \\
residual MLP, radiance metric & 46.3\% & 0.0291\% \\
Mat\'ern head on the flat network's features & 3.82\% & 0.0758\% \\
Mat\'ern head on the weighted network's features & 20.0\% & 0.0282\% \\
per-coordinate combination of the above & \textbf{3.83\%} & \textbf{0.0267\%} \\
\bottomrule
\end{tabular}
\end{table}

Table~\ref{tab:oco2} carries three findings. First, the representation
ladder: the same exact kernel machinery scores $40\%$ on the raw input,
$30\%$ after rescaling each input coordinate by its measured sensitivity (the
metric term of Proposition~\ref{prop:aniso}: the effective rank here is
$17$ of $20$, so the proposition predicts a constant-factor gain and no
change of rate, which is what the scaling experiment shows), and
$3.82\%$ on the trained network's features, past the network's own head. The
raw-input kernel was limited by its features, not by anything kernel-shaped,
and the deep kernel head, an exact solve on learned features, is the
strongest single model on the reduced metric.

Second, the training metric behaves as a budget. The flat-trained and
radiance-trained rows are one architecture and two losses, and each wins the
metric it was trained in by a factor of five or more while losing the other.
The mechanism is visible per coordinate: the radiance metric concentrates on
the leading coefficient, the kernel-flow emulator (whose lengthscales are
tuned per output) predicts that coefficient to $0.0012$ root mean square
against the flat network's $0.0054$, while the flat network is three to four
times more accurate on the trailing thirty coordinates that the radiance
barely sees. Training the network in the radiance metric closes exactly that
gap. A constant-elasticity model of this reallocation is too crude --- the
per-coordinate exponents scatter widely, and coordinate errors do not
decouple because the network shares capacity --- but the direction is robust,
and the practical rule is plain: train in the metric you report.

Third, the regimes of Section~\ref{sec:theory-secmom} are decided by these
numbers before any stack is fit. Against the flat network the raw-input
kernel has $\varrho=0.14$ and $e_n/e_k=0.12$: the condition of
Proposition~\ref{prop:secmom} fails and the kernel is dropped, in contrast to
structural mechanics where the same test keeps it. Because the radiance
metric is diagonal in the reduced coordinates, selecting the best model per
coordinate on the validation split is optimal for both metrics
simultaneously; that per-coordinate combination of the flat and weighted
models is the final surrogate, and it beats the kernel-flow emulator on both
metrics at once ($3.83\%$ against $16.89\%$ reduced, $0.0267\%$ against
$0.0448\%$ radiance). The other bands behave the same way: on WCO2 the deep
kernel heads reach $16.1\%$ and $0.035\%$ against the emulator's $24.1\%$
and $0.060\%$, and on SCO2 the per-coordinate combination reaches $8.0\%$
and $0.043\%$ against $16.1\%$ and $0.115\%$. The code releases the per-band
numbers alongside this paper.
