Throughout this section $\mathcal U=\mathbb R^p$ denotes the (discretized) input
space and $\mathcal V=\mathbb R^q$ the output space, $G:\mathcal U\to\mathcal V$
the target operator, $\mu$ the input distribution, and
$(u_1,v_1),\dots,(u_N,v_N)$ i.i.d.\ draws with $v_n=G(u_n)$; the data carry no
sampling noise in the usual sense, coming from a deterministic solver, though
Section~\ref{sec:diversity} has something to say about solver error. We write
$\ell(w,v)=\|w-v\|_2/\|v\|_2$ for the relative error and
$R(\widehat G)=\mathbb E_{u\sim\mu}\,\ell(\widehat G(u),G(u))$ for the risk, which is
the quantity reported in Section~\ref{sec:experiments}. The results below are
largely elementary, but each one licenses a specific design decision in the
pipeline of Section~\ref{sec:method}, and each has a measurable footprint in
the experiments: one statement per stage, from the symmetry handling and the
kernel-flow term through stacking, the residual correction, its coverage
guarantee, and the choice of nugget. Proofs are collected in
Appendix~\ref{app:proofs}.

\subsection{Symmetry averaging does not increase the risk}
\label{sec:theory-sym}

The elasticity problem behind the benchmark is invariant under the reflection
$x_1\mapsto 1-x_1$: the domain, the boundary partition, and the constitutive
model are all symmetric, the input distribution has a reflection-invariant
covariance, and reflecting the load reflects the stress field. On the discrete
grid this is expressed by two permutation matrices: $S\in\mathbb R^{p\times p}$
reverses the load samples and $T\in\mathbb R^{q\times q}$ reverses the rows of
the stress field. Both are orthogonal involutions. Section~\ref{sec:benchmark}
describes a direct data-driven check of the equivariance
$G(Su)=T\,G(u)$; we treat it as an assumption here.

\begin{proposition}[symmetrization]\label{prop:sym}
Assume $G(Su)=TG(u)$ for $\mu$-a.e.\ $u$, that $\mu$ is invariant under $S$,
and that $T$ is an orthogonal involution ($T^2=I$). For a measurable
$\widehat G:\mathcal U\to\mathcal V$ define its symmetrization
\[
\widehat G_S(u)\;=\;\tfrac12\left(\widehat G(u)+T\,\widehat G(Su)\right).
\]
Then, for every loss $\ell(\cdot,v)$ that is convex in its first argument and
satisfies $\ell(Tw,Tv)=\ell(w,v)$,
\[
\mathbb E_{u\sim\mu}\,\ell\big(\widehat G_S(u),G(u)\big)\;\le\;
\mathbb E_{u\sim\mu}\,\ell\big(\widehat G(u),G(u)\big).
\]
\end{proposition}

The relative error satisfies both hypotheses ($T$ orthogonal gives
$\ell(Tw,Tv)=\ell(w,v)$). Proposition~\ref{prop:sym} justifies the two uses of
the symmetry in the pipeline: averaging each trained model with its reflected
evaluation at test time can only help on average, and training on
reflection-augmented data is ordinary empirical risk minimization for the
symmetrized class. The measured effect is small but strictly positive for
every model we trained (Table~\ref{tab:ablation}).

\subsection{What the kernel-flow regularizer estimates}
\label{sec:theory-kf}

The kernel-flow (KF) loss of Owhadi and Yoo \citep{owhadi2019kernel}, in the
$\ell^2$ variant used by \citet{yoo2021deep} to regularize deep networks,
enters our ablation study as an optional term in the training of the neural
means. Its motivation is often stated informally (``a kernel is good if
half the data predicts the rest''). The following observation makes the
statement exact. Let $z_i=\phi_\theta(u_i)$ be the features of a batch
$B=\{1,\dots,b\}$ ($b$ even), let $k$ be a positive definite kernel on feature
space, and for $C\subset B$, $|C|=b/2$, let $I_C$ denote the $k$-interpolant
of $\{(z_j,v_j):j\in C\}$. The batch KF loss is
\begin{equation}\label{eq:kf}
e_2(\theta;C)\;=\;\sum_{i\in B}\big\|v_i-I_C(z_i)\big\|_2^2 .
\end{equation}

\begin{proposition}[KF loss is a leave-half-out estimate]\label{prop:kf}
Let $C$ be drawn uniformly among the subsets of $B$ of size $b/2$ and assume
$k$ restricted to $\{z_i\}_{i\in B}$ is strictly positive definite. Then
$e_2(\theta;C)=\sum_{i\in B\setminus C}\|v_i-I_C(z_i)\|_2^2$, and
\[
\mathbb E_C\big[e_2(\theta;C)\big]\;=\;\frac b2\;
\mathbb E_{C,\,i\,\sim\,\mathrm{Unif}(B\setminus C)}
\big[\big\|v_i-I_C(z_i)\big\|_2^2\big],
\]
i.e.\ $\tfrac2b\,e_2$ is an unbiased estimator, over the split randomness, of
the expected squared error committed on a held-out point by the kernel
predictor trained on a random half of the batch. When $(B,C)$ are
resampled at every step, the stochastic gradient
$\nabla_\theta e_2(\theta;C)$ is unbiased for
$\nabla_\theta\,\mathbb E_{B,C}[e_2(\theta;C)]$ whenever the expectation and
derivative commute (e.g.\ under local Lipschitz domination).
\end{proposition}

The training loss we minimize for a neural mean is a sum of an empirical risk
term and, optionally, $\beta\,e_2$: the first term is a linear functional of
the empirical measure of the batch and measures fit, while
Proposition~\ref{prop:kf} shows the second measures the \emph{generalization
of a kernel method built on the learned features}. This is the sense in which
a KF-regularized network is trained to be a good feature map for the kernel
correction that follows.

\subsection{Stacking on a held-out split is safe}
\label{sec:theory-stack}

Between the means and the correction sits one more fitted object: the convex
weights of the stacked ensemble, chosen to minimize the empirical risk on the
validation split. Two elementary facts justify this step. First, by convexity,
a convex combination of predictors is never worse than the corresponding
weighted average of their risks, so stacking cannot be hurt by a weak member
beyond its weight. Second, the weights live on a low-dimensional simplex and
are fitted on hundreds of samples, so the selection cannot meaningfully
overfit; the following proposition quantifies this with explicit constants.

\begin{proposition}[validation stacking]\label{prop:stack}
Let $f_1,\dots,f_M:\mathcal U\to\mathcal V$ be fixed maps, let
$\Delta=\{w\in\mathbb R^M:w_m\ge0,\sum_m w_m=1\}$, and for $w\in\Delta$ write
$F_w=\sum_m w_m f_m$. Assume the members' per-sample relative errors are
bounded: $\ell(f_m(u),G(u))\le B$ for $\mu$-a.e.\ $u$ and every $m$.
\begin{enumerate}
\item[(i)] For every $w\in\Delta$,
$\ell(F_w(u),G(u))\le\sum_m w_m\,\ell(f_m(u),G(u))$ pointwise; in particular
$R(F_w)\le\sum_m w_m R(f_m)\le\max_m R(f_m)$, and $\ell(F_w(u),G(u))\le B$.
\item[(ii)] Let $\widehat w$ minimize the empirical risk
$\widehat R(w)=\tfrac1m\sum_{i=1}^m\ell(F_w(\tilde u_i),G(\tilde u_i))$ over
$\Delta$, computed on a validation sample
$\tilde u_1,\dots,\tilde u_m\sim\mu$ independent of $f_1,\dots,f_M$. Then for
every $\delta\in(0,1)$, with probability at least $1-\delta$,
\[
R(F_{\widehat w})\;\le\;\min_{w\in\Delta}R(F_w)\;+\;\frac{8B}{m}\;+\;
B\sqrt{\frac{2\big((M-1)\log\!\big(m(M-1)+1\big)+\log(2/\delta)\big)}{m}} .
\]
\end{enumerate}
\end{proposition}

With $M=4$ members, $m=1000$ validation samples and $\delta=0.05$ the right-hand
excess evaluates to about $0.24\,B$; at $B=0.25$, the largest per-sample member
error we observe, this caps the selection cost at roughly six percentage
points. The bound is conservative, as such bounds are: the realized
validation-to-test gap of the stack in Section~\ref{sec:experiments} is two
orders of magnitude smaller. Its role is the scaling $\sqrt{M\log m/m}$, which
says that fitting a handful of convex weights on a thousand held-out samples
is statistically almost free; this is why the pipeline spends its validation
data there and on the kernel hyperparameters and nowhere else.

\subsection{What the stack can achieve}
\label{sec:theory-secmom}

Proposition~\ref{prop:stack} says fitting the weights is safe; it does not say
mixing is worth anything. That is decided by a single measurable matrix. For a
map $f$ write the \emph{normalized residual} at $u$ as
$\rho_f(u)=\big(f(u)-G(u)\big)/\|G(u)\|_2$, so that
$\ell(f(u),G(u))=\|\rho_f(u)\|_2$.

\begin{proposition}[second-moment identity for convex stacks]\label{prop:secmom}
Let $f_1,\dots,f_M$ be fixed maps and define
$S\in\mathbb R^{M\times M}$ by
$S_{mk}=\mathbb E_{u\sim\mu}\,\langle\rho_{f_m}(u),\rho_{f_k}(u)\rangle$.
For $w$ in the simplex $\Delta^{M-1}$ let $f_w=\sum_m w_mf_m$. Then
\[
\mathbb E_{u\sim\mu}\,\ell\big(f_w(u),G(u)\big)^2\;=\;w^\top S\,w ,
\]
so the squared-metric risk of every convex stack is determined by $S$, and
the best achievable is $\min_{w\in\Delta^{M-1}}w^\top Sw$. In particular:
\begin{enumerate}
\item[(i)] (two members) if $S$ has diagonal $(e_1^2,e_2^2)$ with $e_1\le e_2$
and correlation $\varrho=S_{12}/(e_1e_2)$, mixing in the weaker member
strictly helps if and only if $\varrho<e_1/e_2$, in which case the optimum is
interior with value $e_1^2e_2^2(1-\varrho^2)/(e_1^2+e_2^2-2\varrho e_1e_2)$;
\item[(ii)] (equicorrelated family) if $S_{mm}=e^2$ and $S_{mk}=\varrho\,e^2$
for all $m\neq k$ with $\varrho\ge0$, then
$\min_{w}w^\top Sw=e^2\big(\varrho+(1-\varrho)/M\big)$, attained at uniform
weights and decreasing to $e^2\varrho$ as $M\to\infty$.
\end{enumerate}
\end{proposition}

Exact equicorrelation never holds in practice, but for convex weights the
floor survives one-sided bounds on the entries of $S$, which are what one
actually measures.

\begin{corollary}[ensembling floor]\label{cor:floor}
If every member satisfies $S_{mm}\ge\bar e^{\,2}$ and every pair satisfies
$S_{mk}\ge\bar\varrho\,\bar e^{\,2}$ with $\bar\varrho\in[0,1]$, then every
convex combination $f_w$ obeys
\[
\mathbb E\,\ell\big(f_w(u),G(u)\big)^2\;=\;w^\top Sw\;\ge\;
\bar e^{\,2}\big(\bar\varrho+(1-\bar\varrho)\,\|w\|_2^2\big)\;\ge\;
\bar e^{\,2}\,\bar\varrho .
\]
No number of additional members with the same error level and mutual
correlations moves the ensemble's root-mean-square relative error below
$\bar e\sqrt{\bar\varrho}$.
\end{corollary}

A second consequence of the same decomposition explains the final OCO-2
surrogate. Evaluation metrics there are diagonal in the output coordinates
(the radiance metric is the weighting $s_z$), and diagonal metrics decouple.

\begin{proposition}[per-coordinate combination under diagonal metrics]
\label{prop:percoord}
For weights $v=(v_{mj})$ that may depend on the output coordinate, let
$f_v(u)_j=\sum_m v_{mj}f_m(u)_j$. For any diagonal metric with positive
weights $w$,
\[
\mathbb E\big\|\mathrm{diag}(w)\big(f_v(u)-G(u)\big)\big\|_2^2
=\sum_j w_j^2\;\mathbb E\big(f_v(u)_j-G(u)_j\big)^2 ,
\]
so the minimizing weights solve $q$ separate problems, one per output
coordinate, none of which involves $w$. The per-coordinate optimum is the
same for every diagonal metric, and it weakly dominates every combination
whose weights are shared across coordinates, for all such metrics
simultaneously.
\end{proposition}

The practical content: when a problem is scored in several diagonal metrics
at once, the combination stage does not need to know which one matters. Fit
the best combination coordinate by coordinate on validation and the result
serves all of them; only the members need metric-aware training, which is
where the weighted mean of Section~\ref{sec:oco2} enters.

The corollary is the quantity this paper keeps measuring, at two different
levels. Across architecture families on the structural-mechanics benchmark
the pairwise correlations exceed $0.86$ at member errors near $4.7\%$, a
floor of about $4.4\%$, and the corrected stack lands within a tenth of it.
Across random seeds of one architecture on the OCO-2 bands the correlations
are $0.78$, $0.97$ and $0.96$ at member errors of $4.4\%$, $16.4\%$ and
$8.2\%$, floors of $3.9\%$, $16.2\%$ and $8.0\%$, and the measured
combinations sit on all three. When an ensemble is at its floor the
corollary says where improvement cannot come from; on OCO-2 it came from
changing the members (the kernel head on learned features) rather than
adding more of them. The reported metric is the mean rather than the root mean
square of $\|\rho_f\|_2$, and the two differ by the dispersion of the
per-sample error; across every pipeline in this paper their ratio stays within
a few percent of $0.93$, so predictions made through $S$ transfer to the
reported metric essentially unchanged. The proposition frames the diversity
experiment of Section~\ref{sec:diversity}: what a new member is worth is
legible in the row it adds to $S$, before any weights are fitted. On this
benchmark the rows all look alike, pairwise correlations of
\corrNetLo{}--\corrNetHi{} among trained networks of three architecture
families and two losses, \corrKerLo{}--\corrKerHi{} against the kernel
predictor, so the floor sits just below the best single member, and the
measured stack lands on it. Part (i) is the same
computation specialized to two members; it correctly predicts, from measured
$(e_1,e_2,\varrho)$ alone, which members receive weight in the fitted stack
and which are dropped. Section~\ref{sec:oco2} applies the same test on a
problem where the answer comes out the other way: there the kernel's error is
inflated by a factor the next subsection quantifies, the threshold $e_1/e_2$
collapses, and the kernel is dropped even at a residual correlation of $0.14$.

\subsection{Metric and representation: when a fixed kernel loses to a network}
\label{sec:theory-aniso}

The structural-mechanics benchmark has the neural mean and the isotropic
kernel within half a point of each other. On the OCO-2 emulation problem of
Section~\ref{sec:oco2} the same fixed kernel is an order of magnitude behind
the same network. Two mechanisms can produce such a gap, and they separate
cleanly.

Suppose the operator factors as $G(u)=g(Au)$ with $A\in\mathbb R^{d\times d}$
nonsingular and $g$ of unit $H^{s}$ norm, and that $\mu$ has a density bounded
above and below on a compact domain. Order the singular values
$\sigma_1\ge\dots\ge\sigma_d$ of $A$; they are the rates at which $G$ varies
along the input directions. Suppose $A$ has \emph{approximate rank} $r$ at the
sample size, meaning the trailing singular values fall below the achievable
resolution, $\sigma_{r+1}\lesssim n^{-1/r}$, so the image $A\,\mathcal U$ is
effectively $r$-dimensional at scale $n^{-1/r}$.

\begin{proposition}[isotropic and adapted kernel rates]\label{prop:aniso}
Let $\widehat G^{\mathrm{iso}}_n$ be kernel ridge regression with a Mat\'ern
kernel of smoothness $s$ and a validation-optimal single length scale, and
$\widehat G^{\mathrm{ard}}_n$ the same construction in the metric
$u\mapsto Au$. Up to constants depending on $(\sigma_i)$, $g$, $s$ and the
density bounds, and up to logarithmic factors in $n$, with high probability
over $n$ i.i.d.\ samples,
\[
\mathbb E\big\|G-\widehat G^{\mathrm{iso}}_n\big\|\;\lesssim\;\sigma_1^{\,s}\,n^{-s/d},
\qquad
\mathbb E\big\|G-\widehat G^{\mathrm{ard}}_n\big\|\;\lesssim\;n^{-s/r},
\]
so the ratio of the two bounds is of order $\sigma_1^{\,s}\,n^{\,s(1/r-1/d)}$.
When $A$ is close to full rank ($r\approx d$) the two rates coincide and only
the constant $\sigma_1^{\,s}$ separates them.
\end{proposition}

The proof (Appendix~\ref{app:proofs}) is the scattered-data estimate
$\|G-\widehat G_n\|\lesssim h_X^{\,s}\|G\|_{H^s}$ applied to the two designs:
one length scale must fill all $d$ directions, so $h_X\asymp n^{-1/d}$ and the
norm carries $\sigma_1^{\,s}$; the adapted metric, when $A$ has a spectral gap,
confines the design to $r$ resolved directions and improves the fill distance.
Both displays are upper bounds under the stated idealization --- the exact
metric $A$, a fixed $g$, and the length-scale selection folded into the
logarithmic factors --- and we claim no matching lower bound; the proposition
is used here to calibrate what a metric can and cannot buy, not as a sharp
rate theorem.
The proposition bounds the \emph{metric} part of a network's advantage, the
part an anisotropic kernel would recover, and it says this part is a rate gain
only under an approximate-rank gap; without one it is the constant
$\sigma_1^{\,s}$ alone. The rest is \emph{representational}: a stationary
kernel of fixed smoothness reaches only its native space, at any metric, while
the network adapts its features to the map. The two parts are separated
experimentally by handing the kernel the anisotropic metric and seeing how
much of the gap closes. On the OCO-2 band of Section~\ref{sec:oco2} the input
has effective rank $17$ of $20$, close to full: the proposition predicts the
metric can supply only a constant, not a rate, and the measurement agrees,
$1.4$ of a tenfold gap. The remaining factor is representational, and its
direct evidence is that the same kernel machinery applied to the network's
learned features recovers the network's accuracy and slightly exceeds it.

The representational term has a precise reading through the same
optimal-recovery bound that Section~\ref{sec:theory-or} makes exact. That
bound factors the error as $\|G-m\|_{\mathcal H}\cdot P_\lambda$, a product of
the target's norm in the kernel's native space and a design factor set by the
Gram spectrum. Changing the features changes both in principle, but on the
O2 band only one moves. The effective dimension of the Mat\'ern Gram, the
quantity Lemma~\ref{lem:deff} controls, is nearly identical on the raw input
and on the learned features (at $n=4000$ and a matched median length scale,
$3995$ against $3994$ at a $10^{-8}$ nugget, with equal leading eigenvalue
mass), so the design factor is essentially fixed. The native-space norm of
the target is not: interpolating the same outputs through the two kernels,
the raw-input kernel carries the target at squared norm $1.64\times10^{6}$ and
the feature kernel at $3.87\times10^{4}$, a factor of $42$. The network does
not condition the kernel better; it moves the target to a low-norm corner of
a native space of the same size, where the exact solve reaches it.

This is not an accident of the O2 band; it is what a pulled-back kernel does.
Writing $\varphi$ for the feature map and $k$ for the base kernel, the deep
kernel head uses $k_\varphi(x,x')=k(\varphi(x),\varphi(x'))$.

\begin{proposition}[feature pullback]\label{prop:pullback}
The native space of $k_\varphi$ is $\mathcal H_{k_\varphi}=\{f\circ\varphi:
f\in\mathcal H_k\}$ with
$\|g\|_{\mathcal H_{k_\varphi}}=\min\{\|f\|_{\mathcal H_k}: f\circ\varphi=g
\text{ on }\mathcal X\}$. In particular, if the target factors through the
features as $G=h\circ\varphi$ with $h\in\mathcal H_k$, then
$\|G\|_{\mathcal H_{k_\varphi}}\le\|h\|_{\mathcal H_k}$, and the
optimal-recovery bound of Theorem~\ref{thm:or} for the feature-kernel
regression is governed by $\|h\|_{\mathcal H_k}$ rather than by the norm of
$G$ in the raw-input space.
\end{proposition}

The proof (Appendix~\ref{app:proofs}) is the standard pullback identity for
reproducing kernels. Its content here is the reading: a fixed kernel on the
raw input must carry the whole warped map $G$, whose native-space norm is
large when $G$ has fine structure; the same kernel on the features carries
only the unwarped $h$, and training drives $\varphi$ toward exactly the
factorization that makes $h$ simple. The measured factor of $42$ is that norm
gap, and it is the representational advantage that an anisotropic metric,
which rescales the input but cannot re-express the map, leaves untouched
(recovering only the factor $1.4$).

\subsection{An error bound for the residual kernel correction}
\label{sec:theory-or}

The final stage of the pipeline is kernel ridge regression of the ensemble
residual. The bound below is a vector-valued, residual form of the classical
power-function estimate for kernel interpolation
\citep{wendland2004scattered} and of the optimal-recovery viewpoint of
\citet{owhadi2019operator}; we include the short argument in
Appendix~\ref{app:proofs} to keep the two factors explicit. Fix the mean
$m:\mathcal U\to\mathcal V$ (in the pipeline, the stacked ensemble; in the
statement below $m$ is any fixed map independent of the correction sample) and
write $r=G-m$ for the residual operator with components $r_j$, $j=1,\dots,q$. Let $k$ be a positive definite kernel on $\mathcal U$ with
RKHS $\mathcal H_k$, and assume $r_j\in\mathcal H_k$ for all $j$ with
\[
\|r\|_K^2:=\sum_{j=1}^q\|r_j\|_{\mathcal H_k}^2<\infty .
\]
Given inputs $X=(u_1,\dots,u_n)$, Gram matrix $K=k(X,X)$ and nugget
$\lambda\ge 0$, the correction is
$\widehat r_\lambda(u)=k(u,X)\,(K+n\lambda I)^{-1}R$, $R_{nj}=r_j(u_n)$, and the
corrected predictor is $m+\widehat r_\lambda$. Let
\begin{equation}\label{eq:power}
P_\lambda(u)^2\;=\;k(u,u)-k(u,X)\,(K+n\lambda I)^{-1}k(X,u)
\end{equation}
denote the Gaussian process posterior variance with the same nugget (for
$\lambda=0$ this is the classical power function of the point set $X$).

\begin{theorem}[certified pointwise bound]\label{thm:or}
For every $u\in\mathcal U$ and every $\lambda\ge0$,
\[
\big\|G(u)-m(u)-\widehat r_\lambda(u)\big\|_2\;\le\;\|G-m\|_K\;
\widetilde P_\lambda(u)\;\le\;\|G-m\|_K\;P_\lambda(u),
\]
where
$\widetilde P_\lambda(u)^2=P_\lambda(u)^2-n\lambda\,\|(K+n\lambda
I)^{-1}k(X,u)\|_2^2$.
\end{theorem}

Three comments. First, the inequality is algebraic: it holds pathwise for
every fixed $m$, every dataset, and every $u$, with no probabilistic
assumptions. In the pipeline $m$ is itself fit on the same training set; this
does not affect the validity of the bound, only the reading of
$\|G-m\|_K$ as a random (realized) quantity rather than an a priori one.
Second, the bound factorizes into a quantity that depends only on
the residual operator ($\|G-m\|_K$) and a quantity that depends only on the
kernel and the design ($P_\lambda$), and the second factor is exactly the
posterior standard deviation that a Gaussian process interpretation of the
correction would report. This is the sense in which the bound is certified:
whatever error the corrected surrogate commits at $u$ is controlled by the
reported $P_\lambda(u)$ times a constant that does not depend on $u$. We stress
that this is a statement about a valid upper bound, not about the usefulness of
$P_\lambda$ as a pointwise error \emph{ranking}. Section~\ref{sec:uq} finds
that on this benchmark $P_\lambda$ ranks the error poorly, because the neural
mean supplies most of the error and the relative metric is confounded by output
amplitude; a distribution-free conformal rescaling of $P_\lambda$ nonetheless
attains its nominal coverage on the test set.
Third, the theorem quantifies why one should regress residuals rather than
raw targets: the design factor $P_\lambda$ is the same in both cases, so the
gain of the neural mean is the drop from $\|G-\bar v\|_K$ (kernel-only, mean
$\bar v$) to $\|G-m\|_K$. These norms are not directly observable, but the
RKHS norms of the \emph{fitted} interpolants,
$\smash{\|\widehat r_\lambda\|_K^2=\operatorname{tr}(\alpha^\top K\alpha)}$ with
$\alpha=(K+n\lambda I)^{-1}R$, are computable lower-bound proxies, and they
drop by a factor of about \normRatio{} when the kernel is moved from raw
targets to ensemble residuals (Table~\ref{tab:norms}). The neural mean does
not merely reduce the size of the residual in $L^2$; it leaves behind a
residual operator that is genuinely smoother as seen by the kernel.

\subsection{Distribution-free coverage for the reported band}
\label{sec:theory-conf}

Theorem~\ref{thm:or} controls the error by $\|G-m\|_K\,P_\lambda(u)$, but the
constant $\|G-m\|_K$ is not known a priori, and Section~\ref{sec:uq} shows that
$P_\lambda$ alone ranks the error weakly. What the surrogate reports as
uncertainty is therefore not $P_\lambda$ itself but a rescaling of it,
$q\,P_\lambda(u)$, whose multiplier $q$ is calibrated on the held-out split by
the split-conformal rule
\begin{equation}\label{eq:conf}
q=Q_{1-\alpha}\Big(\big\{s_i:=\|e(\tilde u_i)\|_2/P_\lambda(\tilde u_i)\big\}_{i=1}^m\Big),
\qquad e(u)=G(u)-m(u)-\widehat r_\lambda(u),
\end{equation}
where $Q_{1-\alpha}$ is the $\lceil(1-\alpha)(m+1)\rceil$-th smallest value of
the $m$ validation scores. The point of this construction is that its coverage
needs neither the bound to be tight nor $P_\lambda$ to be a good error ranking;
it needs only exchangeability, which the fixed splits provide.

\begin{proposition}[finite-sample coverage]\label{prop:conf}
Fix the mean $m$, the correction $\widehat r_\lambda$ and the design $X$, and
let $P_\lambda(\cdot)>0$. Suppose the validation and test inputs
$\tilde u_1,\dots,\tilde u_m,u^\star$ are exchangeable (in particular
i.i.d.\ from $\mu$) and drawn independently of $m,\widehat r_\lambda,X$. Let
$q$ be the conformal multiplier \eqref{eq:conf} and define the band
$C(u)=\{v:\|v-m(u)-\widehat r_\lambda(u)\|_2\le q\,P_\lambda(u)\}$. Then
\[
1-\alpha\;\le\;\mathbb P\big(G(u^\star)\in C(u^\star)\big)\;\le\;
1-\alpha+\frac{1}{m+1},
\]
the upper bound holding when the scores are almost surely distinct.
\end{proposition}

The two-sided statement is the standard guarantee of split conformal
prediction \citep{vovk2005algorithmic,lei2018distribution}, transported to the
functional-output setting by taking the nonconformity score to be the relative
residual norm $s=\|e(u)\|_2/P_\lambda(u)$; the proof, a rank argument on the
exchangeable scores, is in Appendix~\ref{app:proofs}. Three points make it the
right closing statement for the uncertainty analysis. It is close to the
quantity we compute: the reported coverage of \uqCover\% at the nominal
$1-\alpha=\uqCoverNom\%$ in Section~\ref{sec:uq} realizes this proposition with
$m=1000$ and $\alpha=0.1$. The $1.6$-point excess over nominal is not the
$1/(m+1)$ slack, which is only a tenth of a point; it is the finite-sample
fluctuation of coverage on a single test set, whose standard deviation
$\sqrt{\alpha(1-\alpha)/m}\approx0.95$ point places \uqCover\% about $1.7$
standard deviations above nominal, well inside the guarantee. One caveat on
the hypotheses: our calibration scores are computed on the same held-out split
used to select the neural-mean checkpoints and tune the correction, so the
exchangeability the proposition assumes is only approximate here; the selection
is low-complexity and the band over-covers, but a calibration split disjoint
from model selection would make the guarantee exact. It is
agnostic to everything the earlier results leave uncertain: the neural mean may
be biased, $P_\lambda$ may correlate with the error weakly or with the wrong
sign, and the bound of Theorem~\ref{thm:or} may be loose, yet the band still
covers at the stated rate. And it is where the uncertainty analysis ends: among
the candidate uncertainty signals we examine, the conformally rescaled
$P_\lambda$ is the only one that carries a guarantee, so it is the one the
surrogate reports.

\subsection{Effective dimension from the Gram spectrum}
\label{sec:theory-rmt}

The remaining design choice is the nugget $\lambda$, selected by
cross-validation in the pipeline. The following exact identity connects that
choice to the spectrum of the Gram matrix and, through it, to random matrix
descriptions of the design.

\begin{lemma}[effective dimension and the empirical Stieltjes transform]
\label{lem:deff}
Let $K$ be a symmetric positive semidefinite $n\times n$ matrix with
eigenvalues $\lambda_1,\dots,\lambda_n\ge0$, let
$\widehat m(z)=\tfrac1n\sum_{i=1}^n(\lambda_i/n-z)^{-1}$ be the Stieltjes
transform of the empirical spectral distribution of $K/n$. Then for every
$\lambda>0$ the effective dimension of ridge regression with nugget $\lambda$
satisfies
\[
d_{\mathrm{eff}}(\lambda)\;:=\;\operatorname{tr}\!\big(K(K+n\lambda I)^{-1}\big)
\;=\;n\big(1-\lambda\,\widehat m(-\lambda)\big).
\]
In particular, if the empirical spectral distribution of $K/n$ converges
weakly to a limit law $F$ as $n\to\infty$, then
$d_{\mathrm{eff}}(\lambda)/n\to 1-\lambda\,m_F(-\lambda)$ with $m_F$ the
Stieltjes transform of $F$.
\end{lemma}

The identity is unconditional; the random-matrix content enters through the
choice of $F$. For the benchmark's simplest reference model, isotropic linear
features, the limit can be evaluated in closed form.

\begin{corollary}[effective dimension under a Mar\v{c}enko--Pastur limit]
\label{cor:mp}
Let $K=XX^\top$ with $X\in\mathbb R^{n\times p}$ having i.i.d.\ entries of
mean zero and variance $\sigma^2$, and let $p/n\to\gamma\in(0,1]$. Then
\[
\frac{d_{\mathrm{eff}}(\lambda)}{n}\;\longrightarrow\;
\gamma\big(1-\lambda\,m(-\lambda)\big),\qquad
m(-\lambda)=\frac{\sqrt{\big(\lambda+\sigma^2(1-\gamma)\big)^2+4\gamma\sigma^2\lambda}
-\big(\lambda+\sigma^2(1-\gamma)\big)}{2\gamma\sigma^2\lambda},
\]
almost surely, where $m$ is the Stieltjes transform of the
Mar\v{c}enko--Pastur law with ratio $\gamma$ and scale $\sigma^2$. In a
simulation with $n=6000$, $p=1800$, $\sigma^2=1.7$, the formula agrees with
the empirical $d_{\mathrm{eff}}(\lambda)/n$ to four decimal places across
$\lambda\in[10^{-3},10]$.
\end{corollary}

Two uses. Quantitatively, the formula makes the nugget--capacity trade-off
explicit for the bulk of a feature Gram matrix: eigenvalue mass of
Mar\v{c}enko--Pastur type is absorbed or discarded by $\lambda$ according to a
single closed-form curve, and outlying (spiked) eigenvalues $x_s$ simply add
their terms $x_s/(x_s+n\lambda)$ on top. Qualitatively, it separates the two
regimes visible in Figure~\ref{fig:spectra}: the Mat\'ern Gram matrix on the
loads is nothing like a Mar\v{c}enko--Pastur bulk (its spectrum decays
exponentially, which is why $d_{\mathrm{eff}}$ is a few hundred out of
19000), whereas the Gram matrices of learned penultimate features do show a
bulk-plus-spikes shape, and for them the corollary describes how much of the
bulk the cross-validated nugget retains. For the feature Gram matrices of the trained networks we
observe the familiar picture of a bulk that is well fitted by a
Mar\v{c}enko--Pastur law \citep{marchenko1967distribution} together with a
small number of outlying eigenvalues carrying the regression signal, as in
spiked covariance models \citep{baik2005phase}; for the Mat\'ern Gram matrix
on the 41-dimensional loads the spectrum decays rapidly and
$d_{\mathrm{eff}}$ is small ($\approx$ a few hundred at the cross-validated
$\lambda$, out of $n=19000$; Figure~\ref{fig:spectra}). This gives a
post-hoc reading of the cross-validated nugget: $\lambda$ lands where
$d_{\mathrm{eff}}(\lambda)$ has absorbed the outlying eigenvalues and the
leading bulk, and further decreasing $\lambda$ buys capacity precisely where
the spectrum carries little signal. It also explains why the exact solve is
affordable: the correction is numerically a problem of dimension
$d_{\mathrm{eff}}$, not $n$.

The effective dimension is also exactly the total budget of the design
factor from Theorem~\ref{thm:or}. Writing $A=K+n\lambda I$, the posterior
variances at the training inputs sum to
\[
\sum_{i=1}^{n}P_\lambda(u_i)^2
=\operatorname{tr}(K)-\operatorname{tr}(K A^{-1}K)
=\operatorname{tr}\!\big(n\lambda\,KA^{-1}\big)
=n\lambda\,d_{\mathrm{eff}}(\lambda),
\]
using $KA^{-1}K=K-n\lambda\,KA^{-1}$. So the same $d_{\mathrm{eff}}$ that reads
the nugget off the spectrum also fixes, up to the factor $n\lambda$, the total
posterior-variance mass the correction distributes over the design; the two
halves of this section are one quantity seen from two sides.
